%===============================================================================
% LaTeX sjabloon voor de bachelorproef toegepaste informatica aan HOGENT
% Meer info op https://github.com/HoGentTIN/latex-hogent-report
%===============================================================================

\documentclass[english,dit,thesis]{hogentreport}

% TODO:
% - If necessary, replace the option `dit`' with your own department!
%   Valid entries are dbo, dbt, dgz, dit, dlo, dog, dsa, soa
% - If you write your thesis in English (remark: only possible after getting
%   explicit approval!), remove the option "dutch," or replace with "english".
\usepackage{xcolor}
\usepackage{listings}
\usepackage{caption} % For proper captions
\usepackage{float} % For H placement specifier
\usepackage{mdframed} % For controlling frames
\usepackage{lipsum} % For blind text, can be removed after adding actual content
\usepackage{placeins}
\usepackage{pifont}                % brings in \ding{}
\usepackage{array}            % for m{width} column
\newcommand{\cmark}{\textcolor{ForestGreen}{\ding{51}}}

% Custom colors for VS Code dark theme
\definecolor{vscodeBackground}{RGB}{30, 30, 30}
\definecolor{vscodeComment}{RGB}{106, 153, 85}
\definecolor{vscodeKeyword}{RGB}{86, 156, 214}
\definecolor{vscodeString}{RGB}{206, 145, 120}
\definecolor{vscodeType}{RGB}{78, 201, 176}
\definecolor{vscodeForeground}{RGB}{220, 220, 220}

% Configure listings for TypeScript code
\lstdefinelanguage{TypeScript}{
  keywords={break, case, catch, continue, debugger, default, delete, do, else, finally, for, function, if, in, instanceof, new, return, switch, this, throw, try, typeof, var, void, while, with, const, let, async, await, class, export, extends, import, super, interface, from},
  keywordstyle=\color{vscodeKeyword},
  ndkeywords={true, false, null, undefined, string, number, boolean, any, void},
  ndkeywordstyle=\color{vscodeType},
  identifierstyle=\color{vscodeForeground},
  sensitive=true,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{vscodeComment},
  stringstyle=\color{vscodeString},
  morestring=[b]',
  morestring=[b]"
}

% Define a completely borderless style for code listings
\definecolor{numberGray}{RGB}{100, 100, 100}
\definecolor{lineNumberBg}{RGB}{40, 40, 40}

% Configure main listing style
\definecolor{frameColor}{RGB}{40, 40, 40}

\lstset{
  backgroundcolor=\color{vscodeBackground},
  basicstyle=\small\ttfamily\color{vscodeForeground},
  frame=none,
  framerule=0pt,
  framesep=0pt,
  xleftmargin=15pt,
  xrightmargin=5pt,
  breaklines=true,
  breakatwhitespace=false,
  postbreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookrightarrow\space}},
  tabsize=2,
  numbers=left,
  numberstyle=\tiny\color{numberGray},
  numbersep=10pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  captionpos=b,
  columns=flexible,
  keepspaces=true,
  aboveskip=1em,
  belowskip=1em
}

% Just use the existing listings setup without trying to redefine environments
\captionsetup[lstlisting]{labelsep=period, labelfont=bf, textfont=it}

% Create a completely borderless frame for listings
\surroundwithmdframed[hidealllines=true, backgroundcolor=vscodeBackground, innerleftmargin=15pt, innerrightmargin=5pt, skipabove=1em, skipbelow=1em]{lstlisting}

% Add a bit of spacing around listings
\lstset{
  aboveskip=1em,
  belowskip=1em
}
% No need to include enumitem package as it's already loaded in hogentreport.cls
% Enable inline lists feature directly after the document class
\makeatletter
\AtBeginDocument{%
  \@ifpackageloaded{enumitem}{%
    \SetEnumitemKey{inline}{itemjoin={,\ },after=.}%
    \newlist{enumerate*}{enumerate*}{3}%
    \setlist[enumerate*]{label=\arabic*.}%
  }{}%
}

%% Pictures to include in the text can be put in the graphics/ folder
\graphicspath{{../graphics/}}

%% For source code highlighting, requires pygments to be installed
%% Compile with the -shell-escape flag!
\usepackage[chapter]{minted}
%% If you compile with the make_thesis.{bat,sh} script, use the following
%% import instead:
% \usepackage[chapter,outputdir=../output]{minted}
\usemintedstyle{solarized-light}

%% Formatting for minted environments.
\setminted{%
    autogobble,
    frame=lines,
    breaklines,
    linenos,
    tabsize=4
}

%% Ensure the list of listings is in the table of contents
\renewcommand\listoflistingscaption{%
    \IfLanguageName{dutch}{Lijst van codefragmenten}{List of listings}
}
\renewcommand\listingscaption{%
    \IfLanguageName{dutch}{Codefragment}{Listing}
}
\renewcommand*\listoflistings{%
    \cleardoublepage\phantomsection\addcontentsline{toc}{chapter}{\listoflistingscaption}%
    \listof{listing}{\listoflistingscaption}%
}

% Other packages not already included can be imported here

%%---------- Document metadata -------------------------------------------------
% TODO: Replace this with your own information
\author{Abdellah El Halimi}
\supervisor{Dhr. A. De Witte}
\cosupervisor{Dhr. C. Dutoict}
% \title[Optionele ondertitel]%
%     {Titel van de bachelorproef}
\title{Accessible Password Management Using Face Recognition for Individuals with Cognitive and Motor Disabilities}
\academicyear{2024-2025}
\examperiod{1}
\degreesought{\IfLanguageName{dutch}{Professionele bachelor in de toegepaste informatica}{Bachelor of applied computer science}}
\partialthesis{false} %% To display 'in partial fulfilment'
%\institution{Internshipcompany BVBA.}

%% Add global exceptions to the hyphenation here
\hyphenation{back-slash}

%% The bibliography (style and settings are  found in hogentthesis.cls)
\addbibresource{bachproef.bib}            %% Bibliography file
\addbibresource{../voorstel/voorstel.bib} %% Bibliography research proposal
\defbibheading{bibempty}{}

%% Prevent empty pages for right-handed chapter starts in twoside mode
\renewcommand{\cleardoublepage}{\clearpage}

\renewcommand{\arraystretch}{1.2}

%% Content starts here.
\begin{document}

%---------- Front matter -------------------------------------------------------

\frontmatter

\hypersetup{pageanchor=false} %% Disable page numbering references

%% Render a Dutch outer title page if the main language is English
% \IfLanguageName{english}{%
%     %% If necessary, information can be changed here
%     \degreesought{Professionele Bachelor toegepaste informatica}%
%     \begin{otherlanguage}{dutch}%
%        \maketitle%
%     \end{otherlanguage}%
% }{}

%% Generates title page content
\maketitle
\hypersetup{pageanchor=true}

\input{voorwoord}
\input{samenvatting}

%---------- Inhoud, lijst figuren, ... -----------------------------------------

\tableofcontents

% In a list of figures, the complete caption will be included. To prevent this,
% ALWAYS add a short description in the caption!
%
%  \caption[short description]{elaborate description}
%
% If you do, only the short description will be used in the list of figures

\listoffigures

% If you included tables and/or source code listings, uncomment the appropriate
% lines.
\listoftables

\listoflistings

% Als je een lijst van afkortingen of termen wil toevoegen, dan hoort die
% hier thuis. Gebruik bijvoorbeeld de ``glossaries'' package.
% https://www.overleaf.com/learn/latex/Glossaries

%---------- Kern ---------------------------------------------------------------

\mainmatter{}

% De eerste hoofdstukken van een bachelorproef zijn meestal een inleiding op
% het onderwerp, literatuurstudie en verantwoording methodologie.
% Aarzel niet om een meer beschrijvende titel aan deze hoofdstukken te geven of
% om bijvoorbeeld de inleiding en/of stand van zaken over meerdere hoofdstukken
% te verspreiden!

\input{inleiding}
\input{standvanzaken}
\input{methodologie}

% Voeg hier je eigen hoofdstukken toe die de ``corpus'' van je bachelorproef
% vormen. De structuur en titels hangen af van je eigen onderzoek. Je kan bv.
% elke fase in je onderzoek in een apart hoofdstuk bespreken.
\input{implementatie}

\input{conclusie}

%---------- Bijlagen -----------------------------------------------------------

\appendix

\chapter{Research Proposal}
The subject of this bachelor's thesis is based on a research proposal that was evaluated beforehand by the promotor. That proposal is included in this appendix.

%% TODO: 
%\section*{Samenvatting}

% Kopieer en plak hier de samenvatting (abstract) van je onderzoeksvoorstel.
    The increasing reliance on online services has elevated the importance of password management in digital security. Traditional authentication methods, such as typing and remembering complex passwords, often pose significant challenges for individuals with cognitive or motor disabilities, such as difficulties with memory recall or typing. These challenges highlight the need for accessible authentication solutions that enhance user satisfaction and experience. This proposal examines the potential of integrating face recognition technology as an accessible authentication mechanism in password management, specifically for individuals with cognitive and motor disabilities.
    The primary research question is: How can face recognition technology improve accessibility in password management for individuals with cognitive and motor disabilities? The research will involve analyzing various face recognition APIs across multiple programming languages to identify the most suitable technology and programming language for this use case. Based on the findings, a web-based prototype password manager will be developed using the chosen technologies. If successful, the next step will involve creating a desktop application to evaluate the scalability and effectiveness of the selected solutions. This iterative methodology ensures a well-informed and comprehensive evaluation of the technologies before full implementation.
    Expected outcomes include improved accessibility and usability compared to traditional password managers, leading to measurable enhancements in user satisfaction and reduced frustration during authentication processes. This study aims to demonstrate the potential of face recognition technology to address accessibility challenges in password management, helping individuals with cognitive and motor disabilities to independently and effectively manage their digital credentials.

% Verwijzing naar het bestand met de inhoud van het onderzoeksvoorstel
\input{../voorstel/voorstel-inhoud}

%%---------- Andere bijlagen --------------------------------------------------
% TODO: Voeg hier eventuele andere bijlagen toe. Bv. als je deze BP voor de
% tweede keer indient, een overzicht van de verbeteringen t.o.v. het origineel.
%\input{...}

%%---------- Backmatter, referentielijst ---------------------------------------

\backmatter{}

\setlength\bibitemsep{2pt} %% Add Some space between the bibliograpy entries
\printbibliography[heading=bibintoc]

\end{document}
