%==============================================================================
% Sjabloon onderzoeksvoorstel bachproef
%==============================================================================
% Gebaseerd op document class `hogent-article'
% zie <https://github.com/HoGentTIN/latex-hogent-article>

% Voor een voorstel in het Engels: voeg de documentclass-optie [english] toe.
% Let op: kan enkel na toestemming van de bachelorproefcoördinator!
\documentclass[english]{hogent-article}

% Invoegen bibliografiebestand
\addbibresource{voorstel.bib}

% Informatie over de opleiding, het vak en soort opdracht

% Voor een voorstel in het Engels, haal de volgende 3 regels uit commentaar
\studyprogramme{Bachelor of applied information technology}
\course{Bachelor thesis}
\assignmenttype{Research proposal}

\academicyear{2024-2025} % TODO: pas het academiejaar aan

% TODO: Werktitel
\title{Accessible Password Management Using Face Recognition for Individuals with Cognitive and Motor Disabilities}

% TODO: Studentnaam en emailadres invullen
\author{Abdellah El Halimi}
\email{abdellah.elhalimimerroun@student.hogent.be}

% TODO: Medestudent
% Gaat het om een bachelorproef in samenwerking met een student in een andere
% opleiding? Geef dan de naam en emailadres hier
% \author{Yasmine Alaoui (naam opleiding)}
% \email{yasmine.alaoui@student.hogent.be}

% TODO: Geef de co-promotor op
\supervisor[Co-promotor]{Dhr. C. Dutoict (Synalco, \href{mailto:"christophe.dutoict@cuarta.be"}{christophe.dutoict@cuarta.be})}

% Binnen welke specialisatierichting uit 3TI situeert dit onderzoek zich?
% Kies uit deze lijst:
%
% - Mobile \& Enterprise development
% - AI \& Data Engineering
% - Functional \& Business Analysis
% - System \& Network Administrator
% - Mainframe Expert
% - Als het onderzoek niet past binnen een van deze domeinen specifieer je deze
%   zelf
%
\specialisation{Mobile \& Enterprise development}
\keywords{Authentication, Face Recognition, Accessibility, Password Manager, Security}

\begin{document}

\begin{abstract}
  The increasing reliance on online services has elevated the importance of password management in digital security. Traditional authentication methods, such as typing and remembering complex passwords, often pose significant challenges for individuals with cognitive or motor disabilities, such as difficulties with memory recall or typing. These challenges highlight the need for accessible authentication solutions that enhance user satisfaction and experience. This proposal examines the potential of integrating face recognition technology as an accessible authentication mechanism in password management, specifically for individuals with cognitive and motor disabilities.
  The primary research question is: How can face recognition technology improve accessibility in password management for individuals with cognitive and motor disabilities? The research will involve analyzing various face recognition APIs across multiple programming languages to identify the most suitable technology and programming language for this use case. Based on the findings, a web-based prototype password manager will be developed using the chosen technologies. If successful, the next step will involve creating a desktop application to evaluate the scalability and effectiveness of the selected solutions. This iterative methodology ensures a well-informed and comprehensive evaluation of the technologies before full implementation.
  Expected outcomes include improved accessibility and usability compared to traditional password managers, leading to measurable enhancements in user satisfaction and reduced frustration during authentication processes. This study aims to demonstrate the potential of face recognition technology to address accessibility challenges in password management, helping individuals with cognitive and motor disabilities to independently and effectively manage their digital credentials.
\end{abstract}

\tableofcontents

% De hoofdtekst van het voorstel zit in een apart bestand, zodat het makkelijk
% kan opgenomen worden in de bijlagen van de bachelorproef zelf.
\input{voorstel-inhoud}

\clearpage

\printbibliography[heading=bibintoc]

\end{document}